### 2021-03-29 outputs/paper/paper.Rmd
repo setup

---
title: "Climate Perceptions in Toronto"
author: "Lorena Almaraz De La Garza"
date: "2021 04 19"
output:
  bookdown::pdf_document2: default
subtitle: "with data from Open Data Toronto"
toc: FALSE
thanks: 'Code and data are available in this GitHub repository: [lalmaraz/toronto_climate_perceptions] (https://github.com/lalmaraz/toronto_climate_perceptions).'
bibliography: references.bib
abstract: X
---

# Introduction

```{r libs, include = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Load libraries
library(bookdown)
library(ggmap)
library(gt)
library(here)
library(httr)
library(janitor)
library(kableExtra)
library(knitr)
library(opendatatoronto)
library(readxl)
packageVersion("readxl")
library(rio)
library(tidyverse)

# Find and get data

questions <- search_packages("climate") %>% 
  list_package_resources() %>% 
  filter(grepl("Readme Data Field Names", name)) %>% 
  get_resource()

raw_data <- search_packages("climate") %>% 
  list_package_resources() %>% 
  filter(grepl("City of Toronto Climate Perceptions SPSS v1", name)) %>% 
  get_resource()

# Save data
write_csv(raw_data,"inputs/data/raw_data.csv")

# Read in data
all_data <- read_csv("inputs/data/raw_data.csv")
```

# Data

# Discussion

## Limitations

## Next Steps

\newpage
# Appendix 1 {#appendix1}
```

\newpage
# References